The natural history of scavenging in vertebrates

### 2016-02-25 natural history of scavenging.tex
added snake data and some bits and pieces from Benbow

@@

Our review is divided up into three sections, namely the land, air and sea. 
These are then subdivided into three geological eras the Cenozoic, Mesozoic and Palaeozoic. 
Each of these environments has a distinct phyiscal character that affects how a species forages for food which has obvious relevance for an animal searching for carrion.  
However, there is some commonality to a scavenger's environment and the problems in finding food that one would encounter. 
Notably, the resource environment of a scavenger is a patchily distributed one, because it is difficult to predict both when and where a carcass is produced. 
As a result of this, any animal existing as a scavenger must maximise its detection capabilities and minimise its locomotory costs \citep{ruxton2004obligate}.
Exploring which groups are likely to have moved towards these traits and thus existed as sccavengers over palaentological time forms the basis of this work.
%remember the journal is an ecological one at heart so I would always keep in mind about making it useful for ecologists over paelo people. In particular from the website "ECOGRAPHY publishes papers focused on broad spatial and temporal patterns, particularly studies of population and community ecology, macroecology, biogeography, and ecological conservation. Studies in ecological genetics and historical ecology are welcomed in the context of explaining contemporary ecological patterns".

@@

@@

\subsection*{Cenozoic}
Birds are the dominant vertebrate fliers today and include the best known scavengers on Earth, the vultures. 
These birds consist of two convergent groups, old world and new world vultures and represent the only example of obligate vertebrate scavengers. 
They have a suite of adaptations that allow them to flourish as obligate scavengers.
Vultures extend the energetic advantage of flight further by engaging in soaring instead of flapping flight, which is even cheaper \citep{hedenstrom1993migration}.
Their efficiency is illustrated by cases of predators like bears and wolves benefiting by taking more carrion in their diet in areas bereft of vultures through competitive release \citep{devault2003scavenging}. 
In flight, birds possess a huge advantage over any terrestrial obligate scavenger. 
Flight affords them the ability to range over a much larger area and detect carrion from an elevated vantage point.
\cite{pennycuick1972soaring} conservatively estimated that a Gyps vulture could identify activity at a carcass 4 km away.

@@

@@

\section{Aquatic Scavengers}
The existence of an obligate scavenger in a marine setting also remains hypothetical \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}. 
Carrion in this environment is produced by dead flesh and marine organisms when their carcasses descend to the sea floor. 
This low-light environment means animals detect resources through chemo- and mechanoreception \citep{ruxton2004energetic}. 
Detection distances are far lower than they would be in the air (< 100 m) as a result. 
However, water is a medium that is conducive to low-cost movement \citep{tucker1975energetic} and so may be able to support a small obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
Although, for the time being this remains conjectural.

>>>>

"Indeed, in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion"\citep{benbow2015introduction}.

@@

Our review is divided up into three sections, namely the land, air and sea. 
These are then subdivided into three geological eras the Cenozoic, Mesozoic and Palaeozoic. 
Each of these environments has a distinct phyiscal character that affects how a species forages for food which has obvious relevance for an animal searching for carrion.  
However, there is some commonality to a scavenger's environment and the problems in finding food that one would encounter. 
Notably, the resource environment of a scavenger is a patchily distributed one, because it is difficult to predict both when and where a carcass is produced. 
As a result of this, any animal existing as a scavenger must maximise its detection capabilities and minimise its locomotory costs \citep{ruxton2004obligate}.
Exploring which groups are likely to have moved towards these traits and thus existed as sccavengers over palaentological time forms the basis of this work.
%remember the journal is an ecological one at heart so I would always keep in mind about making it useful for ecologists over paelo people. In particular from the website "ECOGRAPHY publishes papers focused on broad spatial and temporal patterns, particularly studies of population and community ecology, macroecology, biogeography, and ecological conservation. Studies in ecological genetics and historical ecology are welcomed in the context of explaining contemporary ecological patterns".
"Interkingdom Competition among Vertebrates, Invertebrates, and Microbes"\citep{benbow2015introduction}
"For example, vertebrate scavengers appear to be disadvantaged when humidity and temperature favor microbial and invertebrate reproduction" \citep{benbow2015introduction}

@@

A review by \cite{devault2002scavenging} found occurrences of scavenging spread across five families of snakes and states that this behaviour is "far more common than currently acknowledged."

"\cite{sazima1990necrofagia} predict-ed  that  several  components  of  snake  be-havior contribute to scavenging propensity. They  postulated  that  snakes  that  rely  on chemosensory information to acquire prey might  forage  for  carrion  more  frequently than  those  whose  modes  of  prey  acquisi-tion are driven by visual cues. Additionally, \cite{sazima1990necrofagia} suggested that habitat and diet preferences might be indicators   of   scavenging   propensity   be-cause  they  influence  the  likelihood  of  a particular  species  encountering  carrion naturally. Based on these criteria, they pre-dicted that aquatic or semi-aquatic pisciv-orous  snakes  would  scavenge  more  fre-quently than other species. Moreover, \cite{sazima1990necrofagia}   suggested that  water  currents  that  induce  aggrega-tions of carrion heighten the probability of carrion detection by these snakes (see also \cite{savitzky1992laboratory}).  Also,  chemical  gradients may  be  more  uniform  and  give  more  de-pendable  directional information in water \citep{sazima1990necrofagia}." \citep{devault2002scavenging}

"Investigations of ophidian metabolic re-quirements unveil additional advantages to carrion  utilization.  Snakes  exhibit  exceed-ingly  low  maintenance  metabolisms,  and most  can  survive  on  a  few  scant  feedings per year. It  is, therefore, possible for snakes to rely large-ly  on  infrequent,  less  energy-rich  meals. Carrion, which is by nature ephemeral and unpredictable, may represent such a food source  to  snakes.  Scavenging  might  allow snakes to meet their low metabolic needs more  easily  and  without  the  costs  associ-ated with subduing prey." \citep{devault2002scavenging}

@@

\subsection*{Cenozoic}
Birds are the dominant vertebrate fliers today and include the best known scavengers on Earth, the vultures. 
These birds consist of two convergent groups, old world and new world vultures and represent the only example of obligate vertebrate scavengers. 
They have a suite of adaptations that allow them to flourish in this niche.
Vultures extend the energetic advantage of flight further by engaging in soaring instead of flapping flight, which is even cheaper \citep{hedenstrom1993migration}.
Their efficiency is illustrated by cases of predators like bears and wolves benefiting by taking more carrion in their diet in areas bereft of vultures through competitive release \citep{devault2003scavenging}. 
In flight, birds possess a huge advantage over any terrestrial obligate scavenger. 
Flight affords them the ability to range over a much larger area and detect carrion from an elevated vantage point.
\cite{pennycuick1972soaring} conservatively estimated that a Gyps vulture could identify activity at a carcass 4 km away.

@@

The ability to lower BMR at night is reported in some \textit{Gyps} vultures.

@@

\section{Aquatic Scavengers}
The existence of an obligate scavenger in a marine setting also remains hypothetical \citep{britton1994marine,smith2003ecology,ruxton2004energetic,ruxton2005searching}. 
Carrion in this environment is produced by dead flesh and other marine organisms when their carcasses descend to the sea floor. 
This low-light environment means animals detect resources through chemo- and mechanoreception \citep{ruxton2004energetic}. 
Detection distances are far lower than they would be in the air (< 100 m) as a result. 
However, water is a medium that is conducive to low-cost movement \citep{tucker1975energetic} and so may be able to support a small obligate scavenging fish \citep{ruxton2004energetic,ruxton2005searching}. 
Although, for the time being this remains conjectural.

"although some benthic scavengers (e.g., hagfish: family Myxinidae) rely on necrophagy for a large portion of their diet and may indeed be obligate scavengers" \citep{benbow2015introduction}
